

# R for Data Science and Machine Learning
## Section 6
### Lecture 11: Course notes 
### Lecture 12: Guide to RStudio
### Lecture 13: Introduction to R Basics 
### Lecture 16: Basic data types 

```{r}
# lecture 25: Creating a matrix 

goog <- c(450,451,452,445,468)
msft <- c(230,231,232,233,220)
stocks <- c(goog,msft)
stocks.matrix=matrix(stocks,byrow=T,nrow=2)
stocks.matrix

days <- c('Mon', 'Tue', 'Wed', 'Thu', 'Fri')
st.names <- c("goog","msft")

colnames(stocks.matrix) <- days
rownames(stocks.matrix) <- st.names
print(stocks.matrix)
```

# Section 8
## Lecture 26: Matrix
```{r}
# matrix 
mat <- matrix(1:25,byrow=T,nrow=5)
mat
mat*2
mat/2
mat^2

mat[,c(1,2)]

mat*mat
mat %*% mat
```

## Lecture 27: Matrix operation

```{r}
goog <- c(450,451,452,445,468)
msft <- c(230,231,232,233,220)

stocks <- c(goog,msft)

stocks.matrix=matrix(stocks,byrow=T,nrow=2)
stocks.matrix

days <- c('Mon', 'Tue', 'Wed', 'Thu', 'Fri')
st.names <- c("goog","msft")

colnames(stocks.matrix) <- days
rownames(stocks.matrix) <- st.names
print(stocks.matrix)
```

By using the above stock matrix, we practive matrix operations. 
```{r}
# sum calculation
colSums(stocks.matrix)
rowSums(stocks.matrix)

# row means calculation
rowMeans(stocks.matrix)

# rbind - FB was used as a new row name
FB <- c(111,112,113,120,145)
tech.stocks <- rbind(stocks.matrix,FB)

avg <- rowMeans(tech.stocks)
avg

tech.stocks <- cbind(tech.stocks,avg)
tech.stocks

```

## Lecture28: Matrix selection and indexing
```{r}
mat <- matrix(1:50,byrow=TRUE,nrow=5)
mat[1,]
mat[,]
mat[1:3,]
```

## Lecture29: Factors and categorical matrices

- Nominal categorical - no order
- Ordinal categorical - order

```{r}
animal <- c('d','c','d','c','c')
id <- c(1,2,3,4,5)

# character class -> factor
class(animal)
factor(animal)

# factor 
fact.ani <- factor(animal)
ord.cat <- c('cold','med','hot')
temps <- c('cold','med','hot','hot','hot','cold','med')
temps
summary(temps)

# factor level 
fact_temp <- factor(temps,ordered=TRUE,levels=c('cold','med','hot'))
fact_temp
summary(fact_temp)

```


## Lecture30: Matrix exercises 
```{r}
# R Basic Exercise
#1. two to the power of five
2^5
#2 stock price vector/matrix
stock.prices=c(23,27,23,21,34)
stock.prices
#3 matrix row name
stock.prices=matrix(c(23,27,23,21,34))
rownames(stock.prices)=c("Mon","Tues","Wed","Thu","Fri")
stock.prices
#4 average(mean) computation
mean(stock.prices[,1])
colMeans(stock.prices)
#5. 
over.23 <- stock.prices>23
stock.prices[over.23]

max(stock.prices)
max.price <- stock.prices==max(stock.prices)
max.price

stock.prices[max.price]
```

```{r}
# R Matrix Exercises
# Ex 1: 
require(magrittr)
A <- c(1,2,3) %/% print(A)
B <- c(4,5,6) %/% print(B)

mat <-matrix(A,byrow=T,nrow=1)
matrix <- cbind(mat,B)
matrix <- rbind(mat,B)

rownames(matrix) <- c("A","B")

# Ex 2:
mat <- matrix(1:9,byrow=T,nrow=3)%/% print(mat)

# Ex 3:
is.matrix(mat)

# Ex 4:
mat2 <- matrix(1:25,byrow=T,nrow=5) 
print(mat2)

# Ex 5:
mat2[2:3,2:3]

# Ex 6:
mat2[4:5,4:5]

# Ex 7:
sum(mat2)
sum <- rowSums(mat2)
mat3 <- rbind(mat2,sum)
mat3
sum(mat3[6,])

# Ex 8:
help("runif")
ranmat <- matrix(dnorm(20),nrow=4,ncol=5)
ranmat2 <- matrix(runif(20,min=1,max=100),nrow=4)
ranmat2
```







