---
output:
  word_document: default
  pdf_document: default
  html_document: default
---


# R for Data Science and Machine Learning
## Section 6
### Lecture 11: Course notes 
### Lecture 12: Guide to RStudio
### Lecture 13: Introduction to R Basics 
### Lecture 16: Basic data types 

```{r}
# lecture 25: Creating a matrix 

goog <- c(450,451,452,445,468)
msft <- c(230,231,232,233,220)
stocks <- c(goog,msft)
stocks.matrix=matrix(stocks,byrow=T,nrow=2)
stocks.matrix

days <- c('Mon', 'Tue', 'Wed', 'Thu', 'Fri')
st.names <- c("goog","msft")

colnames(stocks.matrix) <- days
rownames(stocks.matrix) <- st.names
print(stocks.matrix)
```

# Section 8
## Lecture 26: Matrix
```{r}
# matrix 
mat <- matrix(1:25,byrow=T,nrow=5)
mat
mat*2
mat/2
mat^2

mat[,c(1,2)]

mat*mat
mat %*% mat
```

## Lecture 27: Matrix operation

```{r}
goog <- c(450,451,452,445,468)
msft <- c(230,231,232,233,220)

stocks <- c(goog,msft)

stocks.matrix=matrix(stocks,byrow=T,nrow=2)
stocks.matrix

days <- c('Mon', 'Tue', 'Wed', 'Thu', 'Fri')
st.names <- c("goog","msft")

colnames(stocks.matrix) <- days
rownames(stocks.matrix) <- st.names
print(stocks.matrix)
```

By using the above stock matrix, we practive matrix operations. 
```{r}
# sum calculation
colSums(stocks.matrix)
rowSums(stocks.matrix)

# row means calculation
rowMeans(stocks.matrix)

# rbind - FB was used as a new row name
FB <- c(111,112,113,120,145)
tech.stocks <- rbind(stocks.matrix,FB)

avg <- rowMeans(tech.stocks)
avg

tech.stocks <- cbind(tech.stocks,avg)
tech.stocks

```

## Lecture28: Matrix selection and indexing
```{r}
mat <- matrix(1:50,byrow=TRUE,nrow=5)
mat[1,]
mat[,]
mat[1:3,]
```

## Lecture29: Factors and categorical matrices

- Nominal categorical - no order
- Ordinal categorical - order

```{r}
animal <- c('d','c','d','c','c')
id <- c(1,2,3,4,5)

# character class -> factor
class(animal)
factor(animal)

# factor 
fact.ani <- factor(animal)
ord.cat <- c('cold','med','hot')
temps <- c('cold','med','hot','hot','hot','cold','med')
temps
summary(temps)

# factor level 
fact_temp <- factor(temps,ordered=TRUE,levels=c('cold','med','hot'))
fact_temp
summary(fact_temp)

```


## Lecture30: Matrix exercises 
```{r}
# R Basic Exercise
#1. two to the power of five
2^5
#2 stock price vector/matrix
stock.prices=c(23,27,23,21,34)
stock.prices
#3 matrix row name
stock.prices=matrix(c(23,27,23,21,34))
rownames(stock.prices)=c("Mon","Tues","Wed","Thu","Fri")
stock.prices
#4 average(mean) computation
mean(stock.prices[,1])
colMeans(stock.prices)
#5. 
over.23 <- stock.prices>23
stock.prices[over.23]

max(stock.prices)
max.price <- stock.prices==max(stock.prices)
max.price

stock.prices[max.price]
```

```{r}
# R Matrix Exercises
# Ex 1: 
require(magrittr)
A <- c(1,2,3)
B <- c(4,5,6)

mat <-matrix(A,byrow=T,nrow=1)
matrix <- cbind(mat,B)
matrix <- rbind(mat,B)

rownames(matrix) <- c("A","B")

# Ex 2:
mat <- matrix(1:9,byrow=T,nrow=3)
  print(mat)

# Ex 3:
is.matrix(mat)

# Ex 4:
mat2 <- matrix(1:25,byrow=T,nrow=5) 
print(mat2)

# Ex 5:
mat2[2:3,2:3]

# Ex 6:
mat2[4:5,4:5]

# Ex 7:
sum(mat2)
sum <- rowSums(mat2)
mat3 <- rbind(mat2,sum)
mat3
sum(mat3[6,])

# Ex 8:
help("runif")
ranmat <- matrix(dnorm(20),nrow=4,ncol=5)
ranmat2 <- matrix(runif(20,min=1,max=100),nrow=4)
ranmat2
```

# Section 9: Data Frame
## Lecture 33: Data Frame Basics
```{r}
head(state.x77)
class(state.x77)

state.x77.d <- data.frame(state.x77)
head(state.x77.d)
dim(state.x77.d)
summary(state.x77.d)
```

### simple data frame 
```{r}
days <- c("Mon","Tue","Wed","Thur","Fri")
temp <- c(22.2,21,23,24.3,25)
rain <- c(T,T,F,F,T)
df <- data.frame(days,temp,rain)
```

## Lecture34: data frame selection and indexing 
```{r}
# data frame overview
df
df[1,]

# column select by word
df[,'rain']
df[1:5,c('days','temp')]

## column select by $
df$days

# bracket notation
df['days']

# subset function
subset(df,subset=(rain==TRUE))
subset(df,subset=(temp>23.0))
subset(df,subset = temp>23.0)

sorted.temp <- order(df['temp'])
sorted.temp
df[sorted.temp,]

desc.temp <- order(-df$temp)
df[desc.temp,]
```

## LEcture35:Overview of data frame operations (Part1)

```{r}
empty <- data.frame()
empty

c1 <- 1:10

letters
c2 <- letters[1:10]

df <- data.frame(c1,c2)
df <- data.frame(colnames.1=c1,col.names.2=c2)
df
```

### read/write csv
```{r}
# read csv
write.csv(df,file='saved_df.csv')
df2 <- read.csv('saved_df.csv')
df2
```

### Getting information about data frame
```{r}
nrow(df)
ncol(df)
colnames(df)
rownames(df)
str(df)
summary(df)
```

```{r}
# change name of column
colnames(df) <- c("test1","test2")
df$test1

# refer specific row/column
df[[5,2]]
df[[5,'test2']]
df[[5,'test1']] <- 9999
df

df[1,]
```

mpgcars dataset 
```{r}
names(mtcars)
rownames(mtcars)
head(mtcars)

# get vector values in specific row/column
mtcars$mpg
mtcars[,'mpg']
mtcars[,1]

mtcars[['mpg']]

#data frame
mtcars['mpg']
mtcars[c("mpg","cyl")]
```

## Lecture36: Overview of data frame operations (Part2)

```{r}
df2 <- data.frame(test1=2000,test2="new")
df2

require(magrittr)
dfnew <- rbind(df,df2)
print(df)
df$newcol <- 2*df$test1

# copy columns
df$newcol.copy <- df$newcol
print(df)
df[,"newcol.copy2"] <- df$newcol
print(df)

# column names
colnames(df) <- c("1","2","3","4","5")
print(df)
df$`4` <- c(0,0,0,0,0,0,0,0,0,0)

df[1:10,]
df[1:3,]

# data frame head
head(df)
head(df,5)

# exclude specific rows/columns
df[-2,]

head(mtcars)
mtcars[mtcars$mpg>20,]

require(dplyr)
class(mtcars)

# filter / selection functions
mtcars2 <- filter(mtcars, mpg>20&cyl==6) 
select(mtcars,mpg)

mtcars[mtcars$mpg>20 & mtcars$cyl==6,]
mtcars[(mtcars$mpg>20)&(mtcars$cyl==6),c("mpg","cyl","hp")]
subset(mtcars,mpg>20&cyl==6)

head(mtcars)
mtcars[,c(1,2,3)]
is.na(mtcars)

# check any missing point
any(is.na(df))
any(is.na(mtcars$mpg))

df[is.na(df)] <- 0
mtcars$mpg[is.na(mtcars$mpg)] <- mean(mtcars$mpg)
```

## Lecture 37: Data Frames Training Exercises

Ex.1 Recreate data frame
```{r}
Age=c(22,25,26)
weight=c(150,165,120)
Sex=c("M","M","F")
df <- data.frame(Age,weight,Sex)
rownames(df) <- c("Sam","Frank","Amy")
print(df)
```

Ex.2 Check if mtcars is a dataframe using is.data.frame()
```{r}
class(mtcars)
is.data.frame(mtcars)
```

Ex.3 Use as data.frame() to convert a matrix into a dataframe
```{r}
mat <- matrix(1:25,ncol=5)
print(mat)
as.data.frame(mat)
mat
(mat_df <- data.frame(mat)) 
```

Ex.4 Set the built-in data frame mtcars as a variable df. We'll use this df variable for the rest of the excersize.
Ex.5 First 6 rows 
Ex.6 What is the average mpg value for all the cars
Ex.7 select the columns, am, gear and carb
Ex.8 create a new column called performance
Ex.9 2 decimal places by round()
Ex.10 average mgp for cars with more than 100hp and wt value more than 2.5
Ex.11 What is the mpg of the Hornet Sportabout?
```{r}
# Ex.7
(df <- mtcars)
head(df)
mean(mtcars$mpg)

# Ex.8
filter(mtcars, cyl==6)
mtcars[mtcars$cyl==6,]
mtcars[,c("am","gear","carb")]
mtcars$per <- df$hp/df$wt
head(mtcars)

# Ex.9
any(is.na(mtcars$per))
mtcars$per <- round(mtcars$per,2)
head(mtcars)

# Ex.10
mtcars3 <- subset(mtcars,hp>100,wt>2.5)
mean(mtcars3$mpg)

# Ex.11
df
df['Hornet Sportabout',]$mpg
```

# Section 10: List in R
## Lecture 39: List Basics
```{r}
v <- c(1,2,3)
m <- matrix(1:10,nrow=2)
df <- mtcars

class(v)
class(m)
class(df)

my.list <- list(v,m,df)
my.list
my.list[[1]]
class(my.list)

# list assignment with name
my.named.list <- list(sample.vec=v,my.matrix=m,sample.df=df)
my.named.list$sample.vec

double.list <- c(my.named.list,my.named.list)
double.list

str(my.named.list)
```

## Lecture 41: CSV files with R
```{r}
write.csv(mtcars,file="my.example.csv")
ex <- read.csv("my.example.csv")
ex

help(read.csv)
```

## Lecture 42: Excel files with R
```{r}
#install.packages("readxl")
#require(readxl)

# excel load
#excel_sheets('C:/Users/kojiKM.mizumura/Desktop/Sample-Sales-Data.xlsx')
#df <- read_excel('C:/Users/kojiKM.mizumura/Desktop/Sample-Sales-Data.xlsx',sheet="Sheet1")
#head(df)
#sum(df$Value)

#entire.workbook <-lapply(excel_sheets('C:/Users/kojiKM.mizumura/Desktop/Sample-Sales-Data.xlsx'),read_excel,path='C:/Users/kojiKM.mizumura/Desktop/Sample-Sales-Data.xlsx')
#entire.workbook

```

install.packages("xlsx")
library(xlsx)
head(mtcars)
write.xlsx(mtcars,"output_example.xlsx")

## Lecture 43: SQL with R

Connecting R to SQL is actually simple - RODBC library

``` {r}
# install.packages("RODBC")
# library(RODBC)

# myconn <- odbcConnect("Database_Name",uid="User_ID",pwd="password")
# data <- sqlFetch(myconn,"TableName")
# querydat <- sqlQuery(myconn,"SELECT*FROM table")
# close(myconn)
```

- MYSQL: 
The RMySQL package proides an intervace to MySQL
- Oracle:
The ROracle package provides an interface for Oracle
- JDBC: 
The RJDBC package procides access to database throught a JDBC database.

```{r}
# install.packages("RPostgreSQL")
# help(RPostgreSQL)
# ??RPostgreSQL
```


## Lecture 44: Web Scraping with R
HTML and CSS, pipe operator are required for understanding this lecture. 

rvest library
Below is a simple example of using rvest, but the best way to see a good demo of rvest is through the built-in demos by using:
- www.import.io

```{r}
# demo(package="rvest")
# demo(package="rvest",topic="tripadvisor")
```

Now if you are familiar with HTML and CSS a very useful library is rbest. Below we will go ver a simple example from RStudio. Imagine we would like to scrape some information about te Lego Movie from IMDB. We start by downloading and parsing the file with html():

```{r}
# library(rvest)
# lego_movie <- read_html("http://www.imdb.com/title/tt1490017/")
```

To extract the rating, we start with Seletor Gadget to figure out which css selector matches the data we want: strong span.Its the easiest way to determine which sleector extracts the data you are intersed in ). We use html_node to find the first noe that matches that selector, extract its contents with html_text(), and convert it to numeric with as.numeric():
```{r}
#require(magrittr)
# lego_movie %>% 
  # html_node("strong span") %>%
  # html_text() %>% 
  # as.numeric
```

We use a similar process to extract the cast, using html_nodes() to find all nodes that match the selector: 

```{r}
# lego_movie %>%
  # html_nodes("#titleCast.itemprop span") %>%
  # html_text()
```

The titles and authors of recent massage board postings are stored in the third table on the page. We can use html_node() and [[]] to find it, then coerce it to a data frame with html_table():
```{r}
# lego_movie %>%
  # html_nodes("table") %>%
  # .[[3]] %>%
# html_table()
```

# Section 12: R Programming
## Lecture 45: Introduction to programming basics
## Lecture 46: Logical operators 

```{r}
x <- 10
x < 5
x<20 & x>5
T & T

# OR logical operator
(x==10) | (x==100)
```

```{r}
df <- mtcars
df[df$mpg>20,]
df[(df$mpg>20) | (df$hp>100),]
df[(df$mpg>20) & (df$hp>100),]
```

## Lecture 47: If, else, and else if statemnts 

The structure of if statement is: 

if (Condition){
XXX (Code)
} else (condition){
xxx (Code)
}

Indentation is a key for Python. 
```{r}
# example 1
x <- 13
if (x==10){
  # code
  print("X is equal to 10!")
  } else if (x==12) {
  print("X equal to 12")
  }else{
      print("X was not equal to 10 or 12")
  }

# example 2
hot <- FALSE
temp <60

if (temp >80){
  # execute if condition was TRUE
  hot <- TRUE
} else {
  print("Temp is not greater than 80")
}
print(hot)

# example 3
ham <- 10
cheese <- 10
report <- "blank"

if(ham >= 10 & cheese >=10){
  report <- "Strong sales of both ham and cheese"
} else if (ham==0 & cheese==0){
  report <- "No sales today"
} else {
  report <- "we sold something today!"
}
print(report)

```

## Lecture 48: Conditional Statement Exercises

Ex 1.  write a script that will print "Even Number" if the variable is an even number otherwise print "Not Even"
```{r}
x <- 3

if (x %% 2 ==0){
  "Even Number!"
} else {
  "Not Even"
}
```

Ex 2. Write a script that will print "Is a Matrix" if the variable x is a matrix, otherwise print "Not a Matrix".
```{r}
x <- matrix()

if (is.matrix(x)==T){
  "Is a Matrix!"
} else {
  "Not a Matrix"
}
```

Ex 3. Create a script that given a numeric vector x with a length 3, will print out the elements in order from high to low. You must use, if, else if and else statements for logic (This code will be relatively long)
```{r}
x <- c(3,7,1)

if (x[1]>x[2]) {
  # first and second place
  fir <- x[1]
  sec <- x[2]
} else{
  fir <- x[2]
  sec <- x[1]
}

if (x[3] > fir & x[3] >sec){
  third <- sec
  sec <- fir
  fir <- x[3]
} else if (x[3] < fir & x[3] <sec){
  thi <- x[3]
} else {
  third <- sec
  sec <- x[3]
}

print(paste(fir,sec,thi))
```


Ex 4: Write a script that uses if,else if, and else statements to print the max element in a numeric vector with 3 elements.
```{r}
x <- c(20,10,1)
if (x[1]>x[2] & x[1]>x[3]){
  print(x[1])
} else if (x[2]>x[3]){
  print (x[2])
}else {
  print(x[3])
} 
```

## Lecture 50: While Loops 
while (condition){
code executed here 
while condition is true
}
```{r}
x <- 0
while(x<10){
  print(paste0("x is:",x))
  x <- x+1
  if(x==10){
    print("x is now equal to 10! Break Loop!")
    break
    print("Woo I printed too!")
  }
}
```

## Lecture 51: For loops
For loop structure
for (variable in vector){}

```{r}
v <- c(1,2,3,4,5)

for (temp.var in v){
  # execute some code
  # for every temp.var in v
  print("Hello")
  #print(temp.var)
}

for (temp.var in v){
  result <- temp.var+1
  print("The temp.var plus 1 is equal to:")
  print(result)
}

# list operation
my.list <- list(c(1,2,3),mtcars,12)
for (item in my.list){
  print(item)
}

# matrix operation
mat <- matrix(1:25, nrow=5)
print(mat)

for (num in mat){
  print(num)
}  

for (row in 1:nrow(mat)){
  for (col in 1:ncol(mat)){
    print(paste("The selected row is ]",row))
    print(paste("The element at row:",row,"and col:",col,"is",mat[row,col]))  
    }
}
```

## Lecture 52: Functions

```{r}
# Basic sntax of function
name_of_func <- function(inputs1,inputs2,inputs3=45){
  # code execute
  return(result)
}

hello <- function(name){
  print(paste("Hello",name))
}
hello("Sammy")

add_num <- function(num1, num2){
  print(num1+num2)
}
add_num(4,5)

result <- add_num(4,5)
print(result)

# function "times5"
times5 <- function(num){
  my.result <- num*5
  return(my.result)
  #return(num*5)
}
times5(5)

# function "fun"
v <- "I'm a global variable"
stuff <- "I'm global stuff"

fun <- function(stuff){
  print(v)
  stuff <- "Reassign stuff inside of this function fun"
  print(stuff)
}
fun(stuff)


```

## Lecture 53: Functions Training Exercise

Ex. 1: Create a function that will return the product of two integers.
```{r}

prod <- function(num1, num2){
  return(num1*num2)
}
print(prod(3,4))
```

Ex.2: Create a function that accepts two arguments, an integer and a vector of integers. It return TRUE if the integer is present in the vector, otherwise it returns FALSE. Make sure we pay careful attention to our placement of the return(FALSE) line in the function. 
```{r}
num_check <- function (num,v){
  for (ele in v){
    if(ele==num){
      return(TRUE)
    } 
  }
  return(FALSE)
}
num_check(0,c(1,2,3))
```

Ex.3 Create a function that accepts two arguments, an integer and a vector of integers. It returns the count of the number of occurences of the integer in the input vector.
```{r}
num_count <- function(num,v){
  count <- 0
  for (ele in v){
    if (ele ==num){
      count <- count+1
    }
  }
  return(count)
}

print(num_count(2,c(1,0,2,2)))
```

Ex.4 We want to ship bars of alminium. We will create a function that accepts and integer representing the requested kilograms of alminium for the package to be shipped. To fulfill these orders, we have small bars (1 kilogram each). To fullfill these orders, we have small bars (1 kilo each), and big bars (5 kilo each). Return the least number of bar needed. Return FALSE if the required load cannot be created with te bars. 

For example, a load of 6 kg requires a minimum of two bars (1.5 kg bar and 1kg bar). A load of 17 kg requires a minimum of 5 bars (3.5 kg bars and 2.1 kg bars). 

```{r}
bar_count <- function(pack){
  #PAck is the lg load to fill up
  amount.of.ones <- pack %% 5
  amount.of.fives <- (packs - amount.of.ones)/5
  return(amount.of.ones+amount.of.fives)
}
# print(bar_count(17))
```

Ex.5 Create a function that accepts three integer values and returns their sum. However, if an integer value is evenly divisible by 3, then it does not count towards the sum. Return zero if all numbers are evenly divisible by 3. 
Hint: You may want to use the appendix() function.
```{r}
summer <- function(a,b,c){
  out <- c(0)
  if(a%%3 !=0){
    out <- append(a,out)
  }
  if(b%%3 !=0){
    out <- append(b,out)
  }
  if(c%%3 !=0){
    out <- append(c,out)
  }
  return(sum(out))
}
print(summer(1,6,1))
```

Ex.6 Create a function that will return TRUE if an input integer is prime. Otherwise, return FALSE.

```{r}
prime_check <- function(num){
  if(num==2){
    return(TRUE)
  }
  for (x in 2:(num-1)){
    if(num%%x==0){
      return(FALSE)
    }
  }
return(TRUE)
}
```

# Section 13: Advanced R Programming 
## Lecture 56: Built-in R Features

- seq(): create sequences
- sort(): sort a vecotr
- rev(): revser elements in object
- append(): merge objects together (works on vectors/lists)

### seq() function
```{r}
seq(0,10,by=2)
seq(0,100,by=10)
```

### sort() function
sort order of vector elements 
```{r}
v <- c(1,4,7,2,13,3,11)
v
sort(v)
sort(v,decreasing=TRUE)

cv <- c("b","d","a","C")
sort(cv)
```

### rev() function
```{r}
rev(v)

# str
str(v)
str(mtcars)

```

### append()
```{r}
v <- 1:10
v2 <- 35:40

append(v,v2)
```

There are several useful functions for data type.
- is. check data type
- as.: convert

### is.()
```{r}
is.vector(v)
is.data.frame(v)
```

### as.()
```{r}
as.list(v)
is.list(v)

v1 <- as.matrix(v)
```

## Lecture 57: Apply() function

sample numbers
```{r}
print(sample(x=1:100,3))
v <- c(1,2,3,4,5)

addrand <- function(x){
  ran <- sample(1:100,1)
  return(x+ran)
}
addrand(10)
```

### lapply
The result is returned in a list. 
```{r}
#lapply - list apply
(result <- lapply(v,addrand))
```

### sapply
vector/matrix return
```{r}
(result <- sapply(v,addrand))
?apply
```

Another example is as follows. 
```{r}
time2 <- function(num){
  return(num*2)
}

require(magrittr)
(result <- sapply(v,time2))
```

Anonymous function is useful to create function efficiently, and for complex comuputation tasks.
- similar to lambda in python
```{r}

# anonymous functions
## formal
addrand <- function(){
  #code 
}

## anonymous expression
times2 <- function(num){num*2}
result <- sapply(v,function(num){num*2})
print(result)
```

### apply with multiple imputs 

```{r}
v <- 1:5
add_choice <- function(num,choice){
  return(num+choice)
}


print(add_choice(2,10))

# when we want to apply add_choice function to every element

# error code
# sapply(v,add_choice)
# correct code
print(sapply(v,add_choice,choice=100))
```

## Lectore 58: Math Functions with R
A few math functions useful 
- abs(): computes the absolute value
- sum(): returns the sume of all values present in the input
- mean(): computes the arithmetic mean
- round(): rounds values (additional arguments to nearest)

```{r}

2.3331231
round(2.334141411, digits=2)

# R reference card

```

## Lecture 59: Regular Expression
- grepl: returns TRUE/FALSE for a keyword in the data
- grep: count numbers of keyword in the data

```{r}
# example 1
text <- "Hi there, do you know who you are voting for"
text

grepl("voting",text)
grepl("dog",text)
grepl("do you",text)

# example 2
v <- c("a","b","c","d","e")
grep("b",v)
```

## Lecture 60: Data and Timestamps

%d: Day of the month (decimal number)
%m: Month(decimal number)
%b: Moth (abbreviated)
%B: Month (full name)
%y: Year (2 digit)
%Y: Year (4 digit)

```{r}
# today's date
Sys.Date()
today <- Sys.Date()
class(today) #Date

# example 2 - 1990-01-01
c <- "1990-01-01"
class(c)
my.date <- as.Date(c)
class(my.date)

#example 3 - 
my.date <- as.Date("Nov-03-90",format="%b-%d-%y")
my.date

as.Date("June,01,2002",format="%B,%d,%Y")
as.POSIXct("11:02:03",format="%H,%M,%S")

help(strptime)
strptime("date",format="")
```

## Lecture 61: Data Manipulation
- dplyr: manipulate data
- tidyr: clearn data

## Lecture 62: Guide to using Dplyr
We first install packages. 
```{r}
#DPLYR
library(dplyr)
# data load
library(nycflights13)
```
Next, we check nycflights13 data. We are trying serveral useful functions stored in the dplyr package.
- filter() / slice(): select subset of rows in a data frame
- arrange(): 
-

```{r}
??nycflights13
head(flights)
names(flights)
summary(flights)
```

### Filter() function
```{r}
# filter(data, column condition)
(filter(flights,month==11,day==3,carrier=="AA"))
head(flights[flights$month==11 & flights$day==3 & flights$carrier=="AA",])
```

### Slice() function
select rows by position using the dplyr slice package
```{r}
# slice()
slice(flights,1:10)
```

### arrange() function
select and reorder rows
```{r}
head(arrange(flights,year,month,day,arr_time))
head(arrange(flights,year,month,day,desc(arr_time)))
```

###  select() function
to select zoom in specific subsets/columns
```{r}
select(flights,carrier,arr_time,month)
```

### rename() function
```{r}
head(rename(flights,airlines_carrier=carrier))
```

### distinct() function
Distinct returns unique values in the results which is this case all the unique.
```{r}
distinct(select(flights,carrier))
```

### mutate() function
```{r}
require(dplyr)
require(nycflights13)

# mutate - make a new column from existing column
mutate(flights,new_col=arr_delay-dep_delay)

# transmute - only returns created column
head(transmute(flights,new_col=arr_delay*dep_delay))
```

### summarise()
```{r}
summarise(flights, avg_air_time=mean(air_time,na.rm=TRUE))

summarise(flights,total_time=sum(air_time,na.rm=TRUE))
```

## Lecture 64: pipe operator %>%
Example 
```{r}
library(dplyr)
df <- mtcars

# Nesting 
result <- arrange(sample_n(filter(df,mpg>20),size=5),desc(mpg))
print(result)

# better expression
a <- filter(df,mpg>20)
b <- sample_n(a,size=5)
result <- arrange(b,desc(mpg))
print(result)

# pipe operator
# data %>% op1 %>% op2 %>% op3
require(magrittr)

result <- df %>% filter(mpg>20) %>% sample_n(size=5) %>% arrange(desc(mpg))

```

## Lecture 66: Dplyr Training Exercise

Exercise 
- Return rows of cars that have an mpg value greater than 20 and 6 cylinders
- Reorder the data frame by cyl first, and then by descending wt
- select the columns mpg and hp
- select the distinct values of the gear column
- create a new column called "performance" which is calculated by hp divided by wt
- Find the mean mpg vlue using dplyr
- Use pipe operators to get the mean hp value for cars with 6 cylinders
```{r}
require(dplyr)
require(magrittr)

head(mtcars)

#1. filter application
mtcars %>% filter(mpg>20 & cyl==6)
#2. Arrange application
mtcars %>% arrange(cyl, desc(wt))
#3. Select application
select(mtcars,mpg, hp)
#4. distinct application
distinct(select(mtcars,gear))
mtcars %>% select(gear) %>% distinct()
#5. create a new column
mtcars %>% mutate(performance=hp/wt)
#6. summarise
summarise(mtcars,avg.mpg=mean(mpg))
#7. pipe operators 
mtcars %>% filter(cyl==6) %>% summarise(av_hp=mean(hp))
names(mtcars)
```

## Lecture 66: Tidyr package
```{r}
# tidyr
library(tidyr)
# data.table package
library("data.table")
```

1. gather()
collapse multiple columns to keep their values
```{r}
comp <- c(1,1,1,2,2,2,3,3,3)
yr <- c(1998,1999,2000,1998,1999,2000,1998,1999,2000)
q1 <- runif(9,min=0,max=100)
q2 <- runif(9,min=0,max=100)
q3 <- runif(9,min=0,max=100)
q4 <- runif(9,min=0,max=100)

df <- data.frame(comp=comp,year=yr,Qtr1=q1,Qtr2=q2,Qtr3=q3,Qtr4=q4)
df

# gather function 
gather(df,Quarter,Revenue,Qtr1:Qtr4)
help("gather")

# pipe operation
library(magrittr)
df %>%  gather(Quater,Revenue,Qtr1:Qtr4)
```

```{r}
stocks <- data.frame(
  time=as.Date("2009-01-01")+0:9,
  x=rnorm(10,0,1),
  y=rnorm(10,0,2),
  z=rnorm(10,0,4))
stocks
stocks.gt <- stocks %>% gather(stock,price,x,y,z)
stocks.gt
```

2. spread()
do opposite of gather() function
```{r}
# spread
stocks.gt %>% spread(stock, price)
```

3. separate()
Turn single column into multiple columns
- non-alphanumeric character
```{r}
library(tidyr)
df <- data.frame(new.col=c(NA,"a.x","b.y","c.z"))
df
# application of separate() function
separate(df,new.col,c('ABC','XYZ'))

df <- data.frame(new.col=c(NA,"a-x","b-y","c-z")) %>% print()
separate(data=df,col=new.col,into=c("ABC","XYZ"),sep="-")
```

4. Unite 
take away multiple colmuns and unite together
```{r}
df.sep=separate(data=df,col=new.col,into=c("ABC","XYZ"),sep="-")
df.sep

# application of unite() function
unite(df.sep,new.joined.col,ABC,XYZ)
# unite(df.sep,new.joined.col,abc,xyz,sep="---")
```

## Lecture 69: ggplot overview 
Main idea is adding layers together 

(1) First three layers
- data
- aesthetics
- geometries
```{r}
library(ggplot2)
pl <- ggplot(mtcars,aes(x=mpg,y=hp))
pl+geom_smooth()+geom_point()
```

(2) Second step - additional layers
- coordinates
- statistics
- facets
- geometries
- aesthetics
- data
```{r}

# facet grid
pl+geom_point()+facet_grid(cyl~.)

# statistic layer: stat_smonnth()
pl+geom_point()+facet_grid(cyl~.)+stat_smooth()


# coordinates
pl2 <- pl + facet_grid(cyl~.)+stat_smooth()
pl2 + coord_cartesian(xlim=c(15,25))

# Theme layer
pl2 + coord_cartesian(xlim=c(15,25))+theme_bw()
```

## Lecture 70: Histograms
```{r}
# package load
library(ggplot2)
library(ggplot2movies)

#movie 
head(movies)
names(movies)

# data & aesthetics
pl <- ggplot(movies,aes(x=rating))
pl+geom_histogram()
pl2 <- pl+geom_histogram(binwidth=0.1,color="red",fill="pink")

# coordinate addition
pl3 <-  pl2 +xlab("Movie Ratings")+ylab("Count")
pl3 + ggtitle("My Title")
```

## Lecture 81: Scatterplots
```{r}
library(ggplot2)
df <- mtcars

# data and aesthetics
head(mtcars)
names(mtcars)
pl <- ggplot(df, aes(x=wt,y=mpg))
(pl + geom_point())

# geometry
pl + geom_point(alpha=0.5,size=3)
pl+geom_point(aes(size=hp,color=hp))
pl+geom_point(aes(shape=factor(cyl),color=factor(hp)),size=5)

# Hex Color code
pl2 <- pl + geom_point(aes(color=hp))
(pl + geom_point(color="#f5de29"))
# color gardient setting 
pl2 + scale_color_gradient(low="blue",high="red")
```

## Lecture 82: Baplots
```{r}
library(ggplot2)
df <- mpg
head(df)

# y axis: count
pl <- ggplot(df,aes(x=class))
pl + geom_bar(aes(color=drv))
pl + geom_bar(aes(fill=drv))
pl + geom_bar(aes(fill=drv),position="dodge")
pl + geom_bar(aes(fill=drv),position="fill")
```

## Lecture 83: Boxplots
```{r}
require(ggplot)
df <- mtcars
ggplot(df,aes(x=cyl,y=mpg))+geom_boxplot()
ggplot(df,aes(x=factor(cyl),y=mpg))+geom_boxplot()
head(df)

# vertical flip
ggplot(df,aes(x=factor(cyl),y=mpg))+geom_boxplot()+coord_flip()
# color
ggplot(df,aes(x=factor(cyl),y=mpg))+geom_boxplot(aes(fill=factor(cyl)))+theme_bw()
```

## Lecture 84: Variable plotting
```{r}
library(ggplot2)
library(ggplot2movies)

# variable plot
pl <- ggplot(movies,aes(x=year,y=rating))
pl+geom_bin2d()

# scale_fill_gradient
pl2 <- pl+geom_bin2d()+scale_fill_gradient(high="red",low="green")
pl2
pl2 + scale_fill_gradient(high="red",low="blue")

# hexbin
library("hexbin")
pl2 <- pl + geom_hex()
pl2
```

## Lecture 85: Coordinate and Facet
```{r}

# coordinate and facet
library(ggplot2)

# ggplot
pl <- ggplot(mpg,aes(x=displ,y=hwy))+geom_point()
pl

# coord cartersian
pl + coord_cartesian(xlim=c(1,4),ylim=c(15,30))
# coord fixed
pl2 <- pl + coord_fixed(ratio=1/3)
print(pl2)

# facet - separated by cyl (four values)
help("facet_grid")
head(mtcars)

# dvided by column dimension
print(pl + facet_grid(.~cyl))
# divided by row dimension
print(pl + facet_grid(drv ~.))
# divided by row/column dimension
print(pl + facet_grid(drv ~ cyl))
```

## Lecture 76: Themes
```{r}
library(ggplot2)

# theme initial set
theme_set(theme_minimal())

pl <- ggplot(mtcars,aes(x=wt,y=mpg))+coord_cartesian(xlim=c(1,5),ylim=c(15,30))
print(pl + theme_dark())

# theme options
install.packages("ggthemes")
require(ggthemes)
print(pl + theme_economist())
```

## Lecture 77: ggplot2 training exercise
```{r}
# dataset: mpg dataset
library(ggplot2)
library(ggthemes)
head(mpg)
```

```{r}
# Histogram
ggplot(mpg,aes(x=hwy))+geom_histogram(fill="orange",color="red",bin=20)

# Barplot
# cyl is not continuous, thsu factor is needed
ggplot(mpg,aes(x=manufacturer))+geom_bar(aes(fill=factor(cyl)))+
  theme_economist()
```

```{r}
# txhousing dataset
head(txhousing)
```

```{r}
plot(txhousing$sales,txhousing$volume)
pl <- ggplot(txhousing,aes(x=sales,y=volume))
pl + geom_point(color="blue",alpha=0.5) + geom_smooth(color="red")
```

## Lecture 79: Data visualization project

For this optional assigment, we well be recreating this plot from the Economist:

we need to download economist assignment data

```{r}
# package load
library(ggplot2)
library(data.table)

# economist data load
# fread by data.table package
df <- fread("Economist_Assignment_Data.csv",drop=1)
head(df)
```

We start drawing a two dimensional plot by ggplot.
```{r}
ggplot(df,aes(x=CPI,y=HDI))+geom_point(aes(color=Region))
pl <- ggplot(df,aes(x=CPI,y=HDI,color=factor(Region)))+geom_point(size=4,shape=1)
pl

# trend line
pl + geom_smooth(aes(group=1))
pl2 <- pl + geom_smooth(aes(group=1),method="lm",formula=y~log(x),se=F,color="red")
```

```{r}
# add text to each plot
pl3 <- pl2 + geom_text(aes(label=Country,check_overlap=TRUE))
print(pl3)

pointsToLabel <- c("Russia", "Venezuela", "Iraq", "Myanmar", "Sudan",
                   "Afghanistan", "Congo", "Greece", "Argentina", "Brazil",
                   "India", "Italy", "China", "South Africa", "Spane",
                   "Botswana", "Cape Verde", "Bhutan", "Rwanda", "France",
                   "United States", "Germany", "Britain", "Barbados", "Norway", "Japan",
                   "New Zealand", "Singapore")

pl3 <- pl2 + geom_text(aes(label = Country), color = "gray20", 
                data = subset(df, Country %in% pointsToLabel),check_overlap = TRUE)
pl3

# facet value 
pl4 <- pl3 + scale_x_continuous(limits=c(0.9,10.5),breaks=1:10)
pl4
```

## LEcture 82: Interactive Visualizations with Plotly
http://plot.ly
```{r}
install.packages("plotly")
```

Plotly:

https://plot.ly

Package to install:

install.packages('plotly')

Reference for gglotly() examples:
https://plot.ly/ggplot2/

Reference for full ggplotly example documentation:
http://ropensci.github.io/plotly-test-table/tables/0e3d5ca144d27d8416318824c1b6ec1421a51045/index.html

```{r}
library(ggplot2)
library(plotly)
pl <- ggplot(mtcars,aes(mpg,wt))+geom_point()
gpl <- ggplotly(pl)
print(gpl)
```

## Lecture 83: Introduction to capstone data project

- H: Hits
- BB: Bases On Balls
- HBP: Hit By Pitch
- AB: At Bats
- SF: Sacrifice Files

```{r batting csv file load}
Batting <- read.csv('Batting.csv')
```

```{r data overview}
# data confimation
print(head(Batting))
str(Batting)
head(Batting$AB)
head(Batting$X2B)
```

### Feature Engineering
We definite two new variables as follows: On Base Percentage and Slugging Percengage. For SLG, we need 1B(singles), which is not in the data frame.However, we can calculate it by H:1B = H-2B-3B-HR
```{r}
# New variable
# Batting Average
Batting$BA <- Batting$H / Batting$AB
tail(Batting$BA)

# On Base Percentage
Batting$OBP <- (Batting$H+Batting$BB+Batting$HBP)/(Batting$AB+Batting$BB+Batting$HBP+Batting$AB + Batting$SF)

# Creating X1B (Singles)
Batting$X1B <- Batting$H - Batting$X2B - Batting$X3B - Batting$HR

# Slugging Percentage (SLG)
Batting$SLG <- ((1*Batting$X1B)+(2*Batting$X2B)+(3*Batting$X3B)+(4*Batting$HR))/Batting$AB
```

Let's check the structure of the data frame using str().
```{r}
str(Batting)
Batting
```

### Mergin Salary Data with Batting Data
We know we don't want the best players, we want the most undervalued platers, meaning we will also need to know current salary information. We have salary information in the csv file. 
```{r Salary csv file load}
Sal <- read.csv("Salaries.csv") 
```

From the summary of each variable in both data frames, minimum yearID does not match (i.e., yearID in Batting is 1871, and 1985 in Salaries).Thus, we use subset to reassign batting only contain data from 1985. 

```{r}
summary(Batting)
summary(Salaries)
```

```{r data clearning}
library(magrittr)
library(dplyr)

# subset() / arrange()
# we exclude yearID less than 1985
Batting <- subset(Batting,yearID>=1985) %>% arrange(playerID)
head(Batting)
summary(Batting)
```

Now its is time to merge  the batting data with the salary data. Since we have players playing multiple years, we will have repetitions of playersIDs for multiple years, meaning we want to merge on both players and years.

Let's use the merge() function to merge the batting and sal data frames by c("playerID", "yearID").
```{r}
combo <- merge(Batting,Sal,by=c("playerID", "yearID"))
summary(combo)
combo
```

### Analyzing the Lost Players
As previously mentioned, the OAkland A's lost 3 key players during the off-season, We'll want to get their stats to see what we have to replace. The players lost were: first baseman 2000 AL MVP Jason Giambi (giambja01) to the New York Yankees, outfielder Johnny Damon (damonkjo01) to the Boston Red Sox and infielder Rainer Gustavo "Ray" Olmedo (saenzol01). 

Use the subset() function to get a data frame called lost_players from those 3 players. Hint: Try to figure out hwo to use %in% to avoid a bunch of or statements.

```{r lost player data subset}
library(magrittr)
lost_players <- combo %>% subset(playerID %in% c("gianbja01","damonjo01","saenzol01")) 
lost_players

library(ggplot2)
ggplot(lost_players,aes(yearID,salary))+geom_point()+facet_grid(playerID~.)
head(lost_players)

# test value calculation

x <- cbind(x1 = rep(3,8), x2 = c(4:1, 2:5))
x
colSums(x) 
```

### Replacement Players
Now we have all the information we need. Here is the final task - Find replacement platyers for the key three players we lost. However, we have three constrains.
1. The total combined salary of the three players can not exceed 15 million dollars.
2.Their combined number of At Bats (AB) needs to be equal to or greater than the lost players.
3.Their mean OBP had to equal to or greater than the mean OBP of the lost players

Use the combo dataframe you previously created as the source of information! Remember to just use the 2001 subset of that dataframe. There's lost of different ways you can do this, so be creative! It should be relatively simple to find 3 players that satisfy the requirements, note that there are many correct combinations available!

```{r filter/data visualization}
# First only grab available players from year 2001
avail.player <- filter(combo,yearID == 2001)

# we make a quick plot to see where I should cut-off for salary in respect to OBP
ggplot(avail.player,aes(x=OBP,y=salary))+geom_point()
```

- There is no point paying above 8 million or so (I'm just eyeballing this number),I'll choose that as a cut off point.
- The total AB of the lost players is 1469, thus, we should cut off our avail.plater at 1500/3 = 500 AB.
- Nowe lets sort by OBP and see what we've got.

Let's grab columns intersted.
```{r key column analysis}
avail.player %>% filter(salary<8000000, OBP>0) %>%
  filter(AB>=500)

possible <- head(arrange(avail.player,desc(OBP)),10)
possible <- possible[,c("playerID","OBP","AB","salary")]
possible

possible <- head(arrange(avail.player,desc(OBP)),10) %>% select("playerID","OBP","AB","salary")
possible
```

